A meta-analysis of fungicide spray schedules to minimise yield loss from powdery mildew in Australian mungbean

### 2020-08-12 03_Preliminary_analysis.Rmd
Update preliminart analysis page

```{r PA_Libraries, include=FALSE, message=FALSE}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse,
               lubridate,
               bomrang,
               openxlsx,
               devtools,
               reshape2,
               lme4,
               kableExtra,
               here)
source(here("R/import_data.R"))
if (!require("theme.usq"))
   devtools::install_github("adamhsparks/theme.usq")
library(theme.usq)
theme_set(theme_usq()
)
```

## Explore and Visualise Data
First let's import the data.
Each row provides the mean yield for each experimental treatment.

```{r echo=TRUE, message=FALSE, warning=FALSE}
PM_MB_means <- import_data() 
```

The data explored in the `PM_MB_means` data frame is a list of mean values from each treatment in each respective experiment.
These preliminary analyses will focus on the change in mungbean yields, which is attributable to the fungicide treatment. 
Later in this document we will also consider the impact of fungicide treatments on disease incidence/severity.
Mean powdery mildew incidence was recorded at the end of each growing season.
Some experiments recorded disease incidence throughout the season and these were summarised with AUDPC.
AUDPC for trials which only provided final powdery mildew incidence was calculated using the record of 'first sign of disease' and final assessment.
More details on how the `PM_MB_means` data was compiled can be found in the ['Data wrangle vingette'](./vignette/191115_PMMB_DataWrangling_PM.html)

@@

   8 In-crop rainfall

   9 Irrigated versus non-irrigated trials

   10 Mean daily temperature in the weeks following spray application

   11 Final disease rating

   12 Two pathogen species cause PM (new discovery in 2018)

The data plotted below are from `r length(unique(PM_MB_means$trial_ref))` field trials between (`r first(levels(PM_MB_means$year))` - `r last(levels(PM_MB_means$year))`) of which the details are described in the [studies considered in meta-analysis section](#studies-considered-in-meta-analysis).

@@

First let's standardise the sulphur treatments.
Some early experiments tested different formulations.
There was no difference in the sulphur formulations so we will refer to all of them as "sulphur".

```{r sulphur_simplify}
# Remove the variation in different sulphur formulations
unique(PM_MB_means[grep("sulphur", PM_MB_means$fungicide_ai),
                   "fungicide_ai"])
PM_MB_means[grep("sulphur", PM_MB_means$fungicide_ai), "fungicide_ai"] <-
   "sulphur"
```

Let's explore how many different types of fungicide were used and at what frequency.

```{r Fungicides}
PM_MB_means %>%
   group_by(fungicide_ai, trial_ref) %>%
   summarise() %>%
   count(sort = TRUE) %>%
   rename(Trials = n) %>%
   ggplot(aes(x = reorder(fungicide_ai, Trials), y = Trials)) +
   xlab("Fungicide active ingredient") +
   ylab("N Trials") +
   geom_col() +
   scale_fill_usq() +
   ggtitle(label = "Number of trials in which the\nspecified fungicide was used") +
   scale_colour_usq() +
   coord_flip()
```

The demethylation inhibitors (DMI), tebuconazole and propiconazole, are used in the highest frequencies, along with sulphur.
The DMIs have the same fungicide mode of action and are good candidates to be pooled in the meta-analysis.

Amistar Xtra and Custodia both contain strobilurin and triazole, however, because they contain differing dose ratios (inverted) pooling may not be appropriate.

Perhaps best way forward is to focus the meta-analysis on only the DMIs.
This can then be compared to an additional meta-analysis including azoxystrobin as a comparison.

@@

```{r Calculate_yield-savings}
PM_MB_means$grain_yield.t.ha. <-
   as.numeric(PM_MB_means$grain_yield.t.ha.) # NAs produced due to some cells having text description to why there is no specific data

for (i in unique(PM_MB_means$trial_ref)) {
   # Loop code over each trial reference
   dat1 <- PM_MB_means[PM_MB_means$trial_ref == i, ]

@@

      # Reassign values to row_spacing subset
      dat1[dat1$row_spacing == j, ] <- dat2
   }
   # Reassign values to trial_ref subset
   PM_MB_means[PM_MB_means$trial_ref == i, ] <- dat1
}

# ____________________________________________
# calculate yield gain and prop_yield gain for experiments with no row spacing data
dat1 <- PM_MB_means[PM_MB_means$trial_ref == "AM1303", ]

@@

dat1$yield_gain <- dat1$grain_yield.t.ha. - controlY
dat1$prop_YG <- dat1$yield_gain / controlY
dat1[dat1$fungicide_ai == "control", c("yield_gain", "prop_YG")] <-
   c(NA, NA)
PM_MB_means[PM_MB_means$trial_ref == "AM1303", ] <- dat1

write.csv(PM_MB_means,
          "cache/1911_PM_MB_means&Ygains.csv")
```

@@

```{r Fungicide_x_nSprays_tble2, message=FALSE}
PM_MB_means %>%
   group_by(fungicide_ai, total_fungicide) %>%
   summarise(
      n = length(prop_YG),
      lower_2.5 = quantile(prop_YG, na.rm = T, c(0.025)),
      median = median(prop_YG, na.rm = T),
      mean = mean(prop_YG, na.rm = T),
      upper_97.5 = quantile(prop_YG, na.rm = T, c(0.975))
   ) %>%
   filter(n >= 5) %>% # remove any fungicide groups with less than 5 observations
   arrange(desc(median)) %>%
   kable(
      caption = "Fungicide effect on proportion of yield gain.",
      col.names = c(
         "Fungicide active ingredient",
         "Total fungicide applications",
         "n Treatments",
         "2.5% quartile",
         "Median proportional grain yield",
         "Mean proportional grain yield",
         "97.5 % quartile"
      ),
      align = "c"
   ) %>%
   kable_styling(fixed_thead = T, full_width = T) %>%
   column_spec(c(2, 3), width = "1cm") %>%
   footnote(general = "Fungicides with less than five observations were omitted from this table")
```

**  **  *

@@

```{r Fungicide_x_nSprays_tble1, message=FALSE, warning=FALSE}
PM_MB_means %>%
   mutate(fungicide_ai = as.factor(PM_MB_means$fungicide_ai)) %>%
   filter(total_fungicide == 1 | total_fungicide == 2) %>%
   filter(
      fungicide_ai != "pyrazophos" &
         fungicide_ai != "control" &
         fungicide_ai != "benomyl" &
         fungicide_ai != "Acibenzolar-S-methyl"
   ) %>%
   mutate(fungicide_ai = factor(
      fungicide_ai,
      levels = c(
         "carbendazim",
         "pyraclostrobin",
         "sulphur",
         "200 g/L azoxystrobin + 80 g/L cyproconazole",
         "tebuconazole",
         "propiconazole",
         "200 g/L tebuconazole + 120 g/L azoxystrobin"
      )
   )) %>%
   ggplot(aes(y = prop_YG, x = fungicide_ai)) +
   facet_grid(rows = vars(total_fungicide)) +
   geom_boxplot(aes(fill = fungicide_ai)) +
   geom_hline(aes(yintercept = 0), size = 0.5) +
   labs(y = "Proportion of grain yield saved",
        title = "Grain yield proportion saved\ngrouped by fungicide type") +
   theme(legend.position = "none") +
   scale_fill_usq() +
   coord_flip()
```

There is a small yield effect among the strobilurin fungicide mixes and the DMI fungicides. 
While there is some differences in the variation between tebuconazole and propiconazole, the medians are very similar. 
I think it is safe to combine these fungicides, with the same mode of action in the meta-analysis.
The follow-up spray seems to show an increase in yield protection.
Interestingly sulphur sprays show good protection when sprayed twice.

@@

```{r tebuAndPropi_dose}
PM_MB_means %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole") %>%
   select(
      trial_ref,
      year,
      location,
      first_sign_disease,
      fungicide_ai,
      dose_ai.ha,
      total_fungicide
   ) %>%
   ggplot(aes(x = as.factor(dose_ai.ha), fill = fungicide_ai)) +
   xlab("Dose (g ai/ha)") +
   ggtitle(label = "Total number of treatments for each respective tebuconazole dose") +
   geom_bar() +
   scale_fill_usq() +
   scale_colour_usq()
```

@@

```{r PropiDoseEffect}
PM_MB_means %>%
   filter(fungicide_ai == "propiconazole") %>%
   ggplot(aes(x = relevel(as.factor(dose_ai.ha), "62.5"), y = prop_YG)) +
   xlab("Dose (g ai/ha)") +
   ggtitle(label = "Yield for each respective propiconazole dose,\nfaceted by number of applications") +
   geom_boxplot(fill = usq_cols("usq charcoal"), alpha = 0.5) +
   facet_grid(cols = vars(total_fungicide))
```

@@

```{r table_PropiDoses}
table(as.character(PM_MB_means[PM_MB_means$fungicide_ai == "propiconazole",]$trial_ref),
      PM_MB_means[PM_MB_means$fungicide_ai == "propiconazole",]$dose_ai.ha)
```

Two trials show both doses were applied in the same trial, mung1112/01 and mung1112/02.
Let's look at the difference in proportion of yield saved in these trials.

```{r PropiDoses_TrialCompare}
PM_MB_means[(PM_MB_means$trial_ref == "mung1112/01" |
                PM_MB_means$trial_ref == "mung1112/02") &
                PM_MB_means$fungicide_ai == "propiconazole",] %>%
   mutate(dose_ai.ha = as.factor(dose_ai.ha)) %>%
               ggplot(aes(y = prop_YG, x = dose_ai.ha))+
   geom_boxplot()+
   facet_grid(cols = vars(total_fungicide)) +
   xlab("Dose (g ai/ha)") +
   ggtitle(label = "Yield for each respective propiconazole dose, faceted by number of applications") +
   geom_boxplot(fill = usq_cols("usq charcoal"), alpha = 0.5)
```

This plot is similar to the previous plot, indicating dose needs to be considered in the analysis. Especially as there is a interaction between dose and number of sprays.

@@

Let's look at the frequency of sprays per fungicide.  
```{r Fungicide_x_nSprays}
table(PM_MB_means$fungicide_ai, PM_MB_means$total_fungicide)
```

@@

```{r spray_relative2FS}
PM_MB_means %>%
   mutate(
      s1_DfromFS = fungicide_application_1 - first_sign_disease,
      s2_DfromFS = fungicide_application_2 - first_sign_disease,
      s3_DfromFS = fungicide_application_3 - first_sign_disease,
      s4_DfromFS = fungicide_application_4 - first_sign_disease,
      s5_DfromFS = fungicide_application_5 - first_sign_disease
   ) %>%
   gather(
      key = spray,
      value = n_days,
      s1_DfromFS,
      s2_DfromFS,
      s3_DfromFS,
      s4_DfromFS,
      s5_DfromFS
   ) %>%
   drop_na(n_days) %>%
   ggplot(aes(x = spray, y = as.numeric(n_days))) +
   geom_violin() +
   scale_x_discrete(labels = c(
      "First spray",
      "Second spray",
      "Third spray",
      "Fourth spray",
      "Fifth spray"
   )) +
   ylab("Days to First Sign of Disease") +
   facet_grid(cols = vars(total_fungicide)) +
   coord_flip()
```

The plot shows single fungicide applications are mostly made at zero, first sign of disease in the crop.
Experiments that tested two spray treatments also tend to make the first application at first sign of disease, and the second spray between 10 and 25 days later.
Three spray treatments seemed to be mostly used in experiments where preventative sprays as the first fungicide application were made.

@@

```{r Host_genotype}
PM_MB_means %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole") %>%
   group_by(host_genotype, fungicide_ai, trial_ref) %>%
   summarise() %>%
   count() %>%
   rename(Treatments = n) %>%
   ggplot(aes(x = host_genotype, y = Treatments, fill = fungicide_ai)) +
   xlab("Cultivar") +
   ylab("N Trials") +
   ggtitle(label = "Cultivars used in either tebuconazole or propiconazole trials") +
   geom_col() +
   scale_fill_usq(name = "Fungicide AI")
```

@@

This plot shows a large amount of yield volatility in all cultivars except Green Diamond.
Green Diamond shows little volatility probably because this cultivar was only used in `r length(unique(PM_MB_means[PM_MB_means$host_genotype == "Green Diamond","trial_ref"]))` trial.

@@

```{r row_spacing.m}
PM_MB_means %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole") %>%
   group_by(fungicide_ai, row_spacing, trial_ref) %>%
   summarise() %>%
   count() %>%
   rename(Trials = n) %>%
   ggplot(aes(x = as.factor(row_spacing), y = Trials)) +
   xlab("Row Spacing (m)") +
   ylab("N Trials") +
   ggtitle(label = "Trial row spacing using tebuconazole") +
   geom_col(aes(fill = fungicide_ai),
            position = "dodge") +
   scale_fill_usq(name = "Fungicide AI")
```

@@

```{r RS_severity}
# Which row spacing leads to the higher disease severity
PM_MB_means %>%
   filter(year == 2017 |
             year == 2018) %>%
   filter(fungicide_ai == "control") %>%
   ggplot(aes(y = PM_final_severity, x = factor(row_spacing))) +
   geom_boxplot(fill = usq_cols("usq charcoal"), alpha = 0.5) +
   ylab("Final severity rating") +
   xlab("Row spacing (m)") +
   ggtitle("Powdery mildew variation between different row spacing")
```

```{r RS_severity_x_trial}
# Which row spacing leads to the higher disease severity
PM_MB_means %>%
   filter(year == 2017 |
             year == 2018) %>%
   filter(fungicide_ai == "control") %>%
   ggplot(aes(y = PM_final_severity, x = factor(row_spacing))) +
   geom_boxplot(fill = usq_cols("usq charcoal"), alpha = 0.5) +
   facet_grid(cols = vars(location)) +
   ggtitle("Powdery mildew variation between different row spacing") +
   ylab("Final severity rating") +
   xlab("Row spacing (m)")
```

@@

```{r RS_yield}
# Which row spacing leads to the higher yield potential
PM_MB_means %>%
   filter(year == 2017 |
             year == 2018) %>%
   ggplot(aes(
      y = as.numeric(grain_yield.t.ha.),
      x = as.factor(row_spacing),
      colour = location
   )) +
   geom_jitter(width = 0.1) +
   ggtitle("Grain yield results for different row spacing at three locations") +
   xlab("Row spacing (m)") +
   ylab("Grain yield (t/ha)") +
   scale_colour_usq()
```

@@

```{r first_incidence}
# Create the data frame for the plot
season_dates <- PM_MB_means %>%
   group_by(trial_ref) %>%
   summarise(
      Planting_date = unique(planting_date),
      Harvest_date = unique(harvest_date),
      First_sign_PM = unique(first_sign_disease),
      season_length = as.Date(unique(harvest_date)) - as.Date(unique(planting_date)),
      cropping_season = "Actual"
   )

@@

```{r AUDPC_check}
PM_MB_means %>%
   filter(is.na(AUDPC_m)) %>%
   distinct(trial_ref, location, year,AUDPC_m, first_sign_disease, final_assessment)%>%
   kable(align = "c")
```

The trials `r unique(PM_MB_means[is.na(PM_MB_means$AUDPC_m),"trial_ref"]) ` don't have an estimate of AUDPC and need to be filled in. 
The following script estimates AUDPC from the final disease severity rating and the first sign of disease information.

```{r SingleObsAUDPC}
source(here("R/AUDPC_for_single_obs.R"))
write.csv(PM_MB_means,
          "cache/1911_PM_MB_means&Ygains.csv", row.names = FALSE)
```

@@

plot(grain_yield.t.ha. ~ AUDPC_m, data = PM_MB_means)
abline(lm(grain_yield.t.ha. ~ AUDPC_m, data = PM_MB_means))

@@

```{r AUDPClm}
summary(lm(grain_yield.t.ha. ~ AUDPC_m, data = PM_MB_means))
```

This is not what we would expect and the low R squared value of `r round(summary(lm(grain_yield.t.ha. ~ AUDPC_m, data = PM_MB_means))$adj.r.squared,3)` perhaps there are additional factors influencing this relationship.

@@

```{r AUDPCxTrial}
sm2lg <- PM_MB_means %>%
   group_by(trial_ref) %>%
   summarise(TY = mean(grain_yield.t.ha., na.rm = TRUE)) %>%
   arrange(TY)

for (i in sm2lg$trial_ref) {
   dat1 <- PM_MB_means[PM_MB_means$trial_ref == i, ]
   if (any(is.na(dat1$grain_yield.t.ha.))) {
      next
   }
   if (any(is.na(dat1$AUDPC_m))) {
      next
   }

   plot(
      grain_yield.t.ha. ~ AUDPC_m,
      data = dat1,
      main = paste(dat1$location[1], (dat1$year[1])),
      ylim = c(0, max(PM_MB_means$grain_yield.t.ha., na.rm = TRUE)),
      xlim = c(0, max(PM_MB_means$AUDPC_m, na.rm = TRUE))
   )
   abline(lm(grain_yield.t.ha. ~ AUDPC_m, data = dat1))
   points(grain_yield.t.ha. ~ AUDPC_m,
          data = dat1[dat1$fungicide_ai == "control",],
          col = "red")

@@

Based on a GRDC report by Sue Thompson [-@SueThompson2016] we want to know what the yield loss was in on cultivar Jade in the 2016 Trials.
```{r}
dat3 <- PM_MB_means %>%
   filter(year == "2016",
          host_genotype == "Jade",
          fungicide_ai == "control")%>%
   group_by(location)%>%
   summarise(C_Yield = mean(grain_yield.t.ha., na.rm = TRUE))

dat3$F_Yield <- PM_MB_means %>%
   filter(year == "2016",
          host_genotype == "Jade",
          fungicide_ai != "control")%>%
   group_by(location)%>%
   summarise(F_Yield = mean(grain_yield.t.ha., na.rm = TRUE)) %>%
   pull(F_Yield)

@@

```{r Yield_vs_AUDPS}
plot(grain_yield.t.ha. ~ AUDPS_m, data = PM_MB_means)
abline(lm(grain_yield.t.ha. ~ AUDPS_m, data = PM_MB_means))

summary(lm(grain_yield.t.ha. ~ AUDPS_m, data = PM_MB_means))

sm2lg <- PM_MB_means %>%
   group_by(trial_ref) %>%
   summarise(TY = mean(grain_yield.t.ha., na.rm = TRUE)) %>%
   arrange(TY)

for (i in sm2lg$trial_ref) {
   dat1 <- PM_MB_means[PM_MB_means$trial_ref == i, ]
   if (any(is.na(dat1$grain_yield.t.ha.))) {
      next
   }
   if (any(is.na(dat1$AUDPS_m))) {
      next
   }

   plot(
      grain_yield.t.ha. ~ AUDPS_m,
      data = dat1,
      main = paste(dat1$location[1], (dat1$year[1])),
      ylim = c(0, max(
         PM_MB_means$grain_yield.t.ha., na.rm = TRUE
      )),
      xlim = c(0, max(PM_MB_means$AUDPS_m, na.rm = TRUE))
   )
   abline(lm(grain_yield.t.ha. ~ AUDPS_m, data = dat1))
   points(grain_yield.t.ha. ~ AUDPS_m,
          data = dat1[dat1$fungicide_ai == "control",],
          col = "red")

>>>>

```{r PA_Libraries, include=FALSE, message=FALSE}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse,
               lubridate,
               bomrang,
               openxlsx,
               devtools,
               reshape2,
               lme4,
               kableExtra,
               here)

if (!require("theme.usq"))
   devtools::install_github("adamhsparks/theme.usq")
library(theme.usq)
theme_set(theme_usq()
)
load("cache/ImportDataAndSelectTrials01.Rdata")
```

## Explore and Visualise Data
Each row provides the mean yield for each experimental treatment.

The data explored in the `PM_MB_dat` data frame is a list of mean values from each treatment in each respective experiment.
These preliminary analyses will focus on the change in mungbean yields, which is attributable to the fungicide treatment. 
Later in this document we will also consider the impact of fungicide treatments on disease incidence/severity.
Mean powdery mildew incidence was recorded at the end of each growing season.
Some experiments recorded disease incidence throughout the season and these were summarised with AUDPC.
AUDPC for trials which only provided final powdery mildew incidence was calculated using the record of 'first sign of disease' and final assessment.

@@

   8 Final disease rating

The data plotted below are from `r length(unique(PM_MB_dat$trial_ref))` field trials between (`r first(levels(PM_MB_dat$year))` - `r last(levels(PM_MB_dat$year))`) of which the details are described in the [Trials considered for inclusion in meta-analysis section](#Trials-considered-for-inclusion-in-meta-analysis).

@@

@@

```{r Calculate_yield-savings}
for (i in unique(PM_MB_dat$trial_ref)) {
   # Loop code over each trial reference
   dat1 <- PM_MB_dat[PM_MB_dat$trial_ref == i, ]

@@

      # Reassign values to row_spacing subset
      dat1[dat1$row_spacing == j, ] <- dat2
   }
   # Reassign values to trial_ref subset
   PM_MB_dat[PM_MB_dat$trial_ref == i, ] <- dat1
}

# ____________________________________________
# calculate yield gain and prop_yield gain for experiments with no row spacing data
dat1 <- PM_MB_dat[PM_MB_dat$trial_ref == "AM1303", ]

@@

dat1$yield_gain <- dat1$grain_yield.t.ha. - controlY
dat1$prop_YG <- dat1$yield_gain / controlY
dat1[dat1$fungicide_ai == "control", c("yield_gain", "prop_YG")] <-
   c(NA, NA)

PM_MB_dat[PM_MB_dat$trial_ref == "AM1303", ] <- dat1
```

@@

```{r Fungicide_x_nSprays_tble2, message=FALSE}
PM_MB_dat %>%
   group_by(fungicide_ai, total_fungicide) %>%
   summarise(
      n = length(prop_YG),
      lower_2.5 = quantile(prop_YG, na.rm = T, c(0.025)),
      median = median(prop_YG, na.rm = T),
      mean = mean(prop_YG, na.rm = T),
      upper_97.5 = quantile(prop_YG, na.rm = T, c(0.975))
   ) %>%
   filter(n >= 5) %>% # remove any fungicide groups with less than 5 observations
   arrange(desc(median)) %>%
   kable(
      caption = "Fungicide effect on proportion of yield gain.",
      col.names = c(
         "Fungicide active ingredient",
         "Total fungicide applications",
         "n Treatments",
         "2.5% quartile",
         "Median proportional grain yield",
         "Mean proportional grain yield",
         "97.5 % quartile"
      ),
      align = "c"
   ) %>%
   kable_styling(fixed_thead = T, full_width = T) %>%
   column_spec(c(2, 3), width = "1cm") %>%
   footnote(general = "Fungicides with less than five observations were omitted from this table")
```

@@

```{r Fungicide_x_nSprays_tble1, message=FALSE, warning=FALSE}
PM_MB_dat %>%
   mutate(fungicide_ai = as.factor(PM_MB_dat$fungicide_ai)) %>%
   filter(total_fungicide == 1 | total_fungicide == 2) %>%
   filter(
      fungicide_ai != "pyrazophos" &
         fungicide_ai != "control" &
         fungicide_ai != "benomyl" &
         fungicide_ai != "Acibenzolar-S-methyl"
   ) %>%
   mutate(fungicide_ai = factor(
      fungicide_ai,
      levels = c(
         "carbendazim",
         "pyraclostrobin",
         "sulphur",
         "200 g/L azoxystrobin + 80 g/L cyproconazole",
         "tebuconazole",
         "propiconazole",
         "200 g/L tebuconazole + 120 g/L azoxystrobin"
      )
   )) %>%
   ggplot(aes(y = prop_YG, x = fungicide_ai)) +
   facet_grid(rows = vars(total_fungicide)) +
   geom_boxplot(aes(fill = fungicide_ai)) +
   geom_hline(aes(yintercept = 0), size = 0.5) +
   labs(y = "Proportion of grain yield saved",
        title = "Grain yield proportion saved\nfaceted by number of sprays") +
   theme(legend.position = "none") +
   scale_fill_usq() +
   coord_flip()
```

There is a small yield effect among the strobilurin fungicide mixes and the DMI fungicides. 
While there is some differences in the variation between tebuconazole and propiconazole, the medians are very similar. 
I think it is safe to combine these fungicides, with the same mode of action in the meta-analysis.
The follow-up spray seems to show an increase in yield protection.

@@

```{r tebuAndPropi_dose}
PM_MB_dat %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole") %>%
   select(
      trial_ref,
      year,
      location,
      first_sign_disease,
      fungicide_ai,
      dose_ai.ha,
      total_fungicide
   ) %>%
   ggplot(aes(x = as.factor(dose_ai.ha), fill = fungicide_ai)) +
   xlab("Dose (g ai/ha)") +
   ggtitle(label = "Total number of treatments for each respective tebuconazole\nor propiconazole dose") +
   geom_bar() +
   scale_fill_usq() +
   scale_colour_usq()
```

@@

```{r PropiDoseEffect}
PM_MB_dat %>%
   filter(fungicide_ai == "propiconazole") %>%
   ggplot(aes(x = relevel(as.factor(dose_ai.ha), "62.5"), y = prop_YG)) +
   xlab("Dose (g ai/ha)") +
   ggtitle(label = "Yield for each respective propiconazole dose,\nfaceted by number of applications") +
   geom_boxplot(fill = usq_cols("usq charcoal"), alpha = 0.5) +
   ylab("Yield gain proportially\nto the no spray control") +
   facet_grid(cols = vars(total_fungicide))
```

@@

```{r table_PropiDoses}
table(as.character(PM_MB_dat[PM_MB_dat$fungicide_ai == "propiconazole",]$trial_ref),
      PM_MB_dat[PM_MB_dat$fungicide_ai == "propiconazole",]$dose_ai.ha)
```

One trial show both doses were applied in the same trial, and mung1112/02.

@@

Let's look at the frequency of sprays per fungicide.  
```{r Fungicide_x_nSprays}
table(PM_MB_dat$fungicide_ai, PM_MB_dat$total_fungicide)
```

@@

```{r spray_relative2FS}
PM_MB_dat %>%
   mutate(
      s1_DfromFS = fungicide_application_1 - first_sign_disease,
      s2_DfromFS = fungicide_application_2 - first_sign_disease,
      s3_DfromFS = fungicide_application_3 - first_sign_disease,
      s4_DfromFS = fungicide_application_4 - first_sign_disease,
      s5_DfromFS = fungicide_application_5 - first_sign_disease
   ) %>%
   gather(
      key = spray,
      value = n_days,
      s1_DfromFS,
      s2_DfromFS,
      s3_DfromFS,
      s4_DfromFS,
      s5_DfromFS
   ) %>%
   drop_na(n_days) %>%
   ggplot(aes(x = spray, y = as.numeric(n_days))) +
   geom_violin() +
   scale_x_discrete(labels = c(
      "First spray",
      "Second spray",
      "Third spray",
      "Fourth spray",
      "Fifth spray"
   )) +
   ylab("Fungicide application timing relative to First Sign of Disease") +
   facet_grid(cols = vars(total_fungicide)) +
   coord_flip()
```

The plot shows single fungicide applications are mostly made at zero, first sign of disease in the crop. Some were made earlier and some around 10 days later.  
Experiments that tested two spray treatments also tend to make the first application at first sign of disease, and the second spray between 10 and 25 days later.
Three spray treatments seemed to be mostly used in experiments where preventative sprays as the first fungicide application were made.

@@

```{r Host_genotype}
PM_MB_dat %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole") %>%
   group_by(host_genotype, fungicide_ai, trial_ref) %>%
   summarise() %>%
   count() %>%
   rename(Treatments = n) %>%
   ggplot(aes(x = host_genotype, y = Treatments, fill = fungicide_ai)) +
   xlab("Cultivar") +
   ylab("N Trials") +
   ggtitle(label = "Cultivars used in either tebuconazole or propiconazole trials") +
   geom_col() +
   scale_fill_usq(name = "Fungicide AI")
```

@@

This plot shows a large amount of yield volatility in all cultivars except Green Diamond.
Green Diamond shows little volatility probably because this cultivar was only used in `r length(unique(PM_MB_dat[PM_MB_dat$host_genotype == "Green Diamond","trial_ref"]))` trial.

@@

```{r row_spacing.m}
PM_MB_dat %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole") %>%
   group_by(fungicide_ai, row_spacing, trial_ref) %>%
   summarise() %>%
   count() %>%
   rename(Trials = n) %>%
   ggplot(aes(x = as.factor(row_spacing), y = Trials)) +
   xlab("Row Spacing (m)") +
   ylab("N Trials") +
   ggtitle(label = "Trial row spacing using tebuconazole") +
   geom_col(aes(fill = fungicide_ai),
            position = "dodge") +
   scale_fill_usq(name = "Fungicide AI")
```

@@

```{r RS_severity}
# Which row spacing leads to the higher disease severity
PM_MB_dat %>%
   filter(year == 2017 |
             year == 2018) %>%
   filter(fungicide_ai == "control") %>%
   ggplot(aes(y = PM_final_severity, x = factor(row_spacing))) +
   geom_boxplot(fill = usq_cols("usq charcoal"), alpha = 0.5) +
   ylab("Final severity rating") +
   xlab("Row spacing (m)") +
   ggtitle("Powdery mildew variation between different row spacing")
```

```{r RS_severity_x_trial}
# Which row spacing leads to the higher disease severity
PM_MB_dat %>%
   filter(year == 2017 |
             year == 2018) %>%
   filter(fungicide_ai == "control") %>%
   ggplot(aes(y = PM_final_severity, x = factor(row_spacing))) +
   geom_boxplot(fill = usq_cols("usq charcoal"), alpha = 0.5) +
   facet_grid(cols = vars(location)) +
   ggtitle("Powdery mildew variation between different row spacing") +
   ylab("Final severity rating") +
   xlab("Row spacing (m)")
```

@@

```{r RS_yield}
# Which row spacing leads to the higher yield potential
PM_MB_dat %>%
   filter(year == 2017 |
             year == 2018) %>%
   ggplot(aes(
      y = as.numeric(grain_yield.t.ha.),
      x = as.factor(row_spacing),
      colour = location
   )) +
   geom_jitter(width = 0.1) +
   ggtitle("Grain yield results for different row spacing at three locations") +
   xlab("Row spacing (m)") +
   ylab("Grain yield (t/ha)") +
   scale_colour_usq()
```

@@

```{r first_incidence}
# Create the data frame for the plot
season_dates <- PM_MB_dat %>%
   group_by(trial_ref) %>%
   summarise(
      Planting_date = unique(planting_date),
      Harvest_date = unique(harvest_date),
      First_sign_PM = unique(first_sign_disease),
      season_length = as.Date(unique(harvest_date)) - as.Date(unique(planting_date)),
      cropping_season = "Actual"
   )

@@

```{r AUDPC_check}
PM_MB_dat %>%
   filter(is.na(AUDPC_m)) %>%
   distinct(trial_ref, location, year,AUDPC_m, first_sign_disease, final_assessment)%>%
   kable(align = "c")
```

The trials `r unique(PM_MB_dat[is.na(PM_MB_dat$AUDPC_m),"trial_ref"]) ` don't have an estimate of AUDPC and need to be filled in. 
The following script estimates AUDPC from the final disease severity rating and the first sign of disease information.

```{r SingleObsAUDPC}
source(here("R/AUDPC_for_single_obs.R"))
write.csv(PM_MB_dat,
          "cache/1911_PM_MB_dat&Ygains.csv", row.names = FALSE)
```

@@

plot(grain_yield.t.ha. ~ AUDPC_m, data = PM_MB_dat)
abline(lm(grain_yield.t.ha. ~ AUDPC_m, data = PM_MB_dat))

@@

```{r AUDPClm}
summary(lm(grain_yield.t.ha. ~ AUDPC_m, data = PM_MB_dat))
```

This is not what we would expect and the low R squared value of `r round(summary(lm(grain_yield.t.ha. ~ AUDPC_m, data = PM_MB_dat))$adj.r.squared,3)` perhaps there are additional factors influencing this relationship.

@@

```{r AUDPCxTrial}
sm2lg <- PM_MB_dat %>%
   group_by(trial_ref) %>%
   summarise(TY = mean(grain_yield.t.ha., na.rm = TRUE)) %>%
   arrange(TY)

for (i in sm2lg$trial_ref) {
   dat1 <- PM_MB_dat[PM_MB_dat$trial_ref == i, ]
   if (any(is.na(dat1$grain_yield.t.ha.))) {
      next
   }
   if (any(is.na(dat1$AUDPC_m))) {
      next
   }

   plot(
      grain_yield.t.ha. ~ AUDPC_m,
      data = dat1,
      main = paste(dat1$location[1], (dat1$year[1])),
      ylim = c(0, max(PM_MB_dat$grain_yield.t.ha., na.rm = TRUE)),
      xlim = c(0, max(PM_MB_dat$AUDPC_m, na.rm = TRUE))
   )
   abline(lm(grain_yield.t.ha. ~ AUDPC_m, data = dat1))
   points(grain_yield.t.ha. ~ AUDPC_m,
          data = dat1[dat1$fungicide_ai == "control",],
          col = "red")

@@

Based on a GRDC report by Sue Thompson [-@SueThompson2016] we want to know what the yield loss was in on cultivar Jade in the 2016 Trials.
```{r}
dat3 <- PM_MB_dat %>%
   filter(year == "2016",
          host_genotype == "Jade",
          fungicide_ai == "control")%>%
   group_by(location)%>%
   summarise(C_Yield = mean(grain_yield.t.ha., na.rm = TRUE))

dat3$F_Yield <- PM_MB_dat %>%
   filter(year == "2016",
          host_genotype == "Jade",
          fungicide_ai != "control")%>%
   group_by(location)%>%
   summarise(F_Yield = mean(grain_yield.t.ha., na.rm = TRUE)) %>%
   pull(F_Yield)

@@

```{r Yield_vs_AUDPS}
plot(grain_yield.t.ha. ~ AUDPS_m, data = PM_MB_dat)
abline(lm(grain_yield.t.ha. ~ AUDPS_m, data = PM_MB_dat))

summary(lm(grain_yield.t.ha. ~ AUDPS_m, data = PM_MB_dat))

sm2lg <- PM_MB_dat %>%
   group_by(trial_ref) %>%
   summarise(TY = mean(grain_yield.t.ha., na.rm = TRUE)) %>%
   arrange(TY)

for (i in sm2lg$trial_ref) {
   dat1 <- PM_MB_dat[PM_MB_dat$trial_ref == i, ]
   if (any(is.na(dat1$grain_yield.t.ha.))) {
      next
   }
   if (any(is.na(dat1$AUDPS_m))) {
      next
   }

   plot(
      grain_yield.t.ha. ~ AUDPS_m,
      data = dat1,
      main = paste(dat1$location[1], (dat1$year[1])),
      ylim = c(0, max(
         PM_MB_dat$grain_yield.t.ha., na.rm = TRUE
      )),
      xlim = c(0, max(PM_MB_dat$AUDPS_m, na.rm = TRUE))
   )
   abline(lm(grain_yield.t.ha. ~ AUDPS_m, data = dat1))
   points(grain_yield.t.ha. ~ AUDPS_m,
          data = dat1[dat1$fungicide_ai == "control",],
          col = "red")
